---
title: 'Data 621 Final Project: Contract Renewal Predictions'
author: Critical Thinking Group 2 - James Topor, Armenoush Aslanian-Persico, Jeff
  Nieman, Scott Karr
date: "July 2016"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    code_folding: show
    theme: spacelab
---

```{r, eval=TRUE, echo=FALSE, include=FALSE}
library(alr3)
library(bestglm)
library(car)
library(ggplot2)
library(dplyr)
library(knitr)
library(lmtest)
library(fBasics)
```


# Part 1: Data Exploration

## Data Dictionary

```{r, echo=FALSE}
url1 <- "https://raw.githubusercontent.com/spsstudent15/2016-02-621-W2/master/621-FP-data-dictionary.csv"

ddict <- read.csv(url1, stringsAsFactors = FALSE)  

kable(ddict)
```

## Summary Statistics

```{r, fig.width = 9, fig.height = 7, include = TRUE, echo=FALSE}
#url <- "https://raw.githubusercontent.com/spsstudent15/2016-02-621-W2/master/621-FP-data-not-transformed.csv"

url <- "https://raw.githubusercontent.com/spsstudent15/2016-02-621-W2/master/621-FP-data-renewals-as-1.csv"




fp1 <- read.csv(url, stringsAsFactors = TRUE)  

attach(fp1)

summary(fp1)

kable(head(fp1))
```

```{r}
#basicStats(fp1)

```


Our data set contains recent service contracts with Company X, some of which were renewed and some which were not renewed. Our goal is to predict whether a service contract will be renewed based on the attributes in the data set. A renewal is considered successful. 

There are 91726 rows of data, each representing a service contract with Company X. 

We have 15 potential predictor variables and one response variable ("RENEWAL_FLAG") that indicates if the contract was renewed.  


## Boxplots and Mosaic Plots
  
```{r, fig.width = 7.5, fig.height = 8.5, echo = FALSE}
# box plots of each predictor variable relative to the response


# par(mfrow=c(4,4)) 

mosaicplot(~ RENEWAL_FLAG + AREA, data = fp1, color = TRUE)

mosaicplot(~ RENEWAL_FLAG + CHANNEL_APPROACH, data = fp1, color = TRUE)

# placeholder CONTRACT_EXPIRATION

mosaicplot(~ RENEWAL_FLAG + CONTRACT_LENGTH, data = fp1, color = TRUE)

mosaicplot(~ RENEWAL_FLAG + SALES_CATEGORY, data = fp1, color = TRUE)

# placeholder SELLER_UNIQUE_ID

mosaicplot(~ RENEWAL_FLAG + CONTRACT_VALUE_CATEGORY, data = fp1, color = TRUE)

mosaicplot(~ RENEWAL_FLAG + CONTRACT_LINE_CATEGORY, data = fp1, color = TRUE)

mosaicplot(~ RENEWAL_FLAG + DISCOUNT_CATEGORY, data = fp1, color = TRUE)

mosaicplot(~ RENEWAL_FLAG + MULTIPLE_SERVICES, data = fp1, color = TRUE)

boxplot(ITEM_COUNT~RENEWAL_FLAG, ylab="ITEM_COUNT", 
        xlab="Renewals", col = "yellow")

# placeholder COST

mosaicplot(~ RENEWAL_FLAG + SALES_STRATEGY, data = fp1, color = TRUE)

```


## Histograms


```{r, fig.width = 9, fig.height = 7, echo = FALSE, eval=FALSE}
#####################################################################
# Make small histograms for each variable

par(mfrow = c(4,3), oma = c(1, 1, 0, 0), mar=c(2, 2, 0, 1) + 2)


hist(AREA, breaks = 30, col = 'yellow')

hist(SERVICE_APPROACH, breaks = 30, col = 'yellow')

hist(CHANNEL_APPROACH, breaks = 30, col = 'yellow')

hist(CONTRACT_EXPIRATION, breaks = 30, col = 'yellow')


hist(CONTRACT_LENGTH, breaks = 30, col = 'yellow')

```



# Part 2: Data Preparation

## Variables to be Transformed


```{r, eval=FALSE}
fp.data <- read.csv("c:/users/hammer/documents/Data Science Docs/CUNY MSDA 621/Summer 2016/Final Project/fin proj.csv", stringsAsFactors = FALSE)

# create a copy of the original data to use for transforming individual variables
fp.t <- fp.data

```



```{r, message = FALSE, warning = FALSE}

# load packages

```

start our data transformation....

1.  Area - go for USA=1 and not USA = 2
```{r, eval=FALSE}
# transform AREA: USA = 1; !USA = 2

fp.t$AREA[which(fp.data$AREA == 'USA')] <- 1
fp.t$AREA[which(fp.data$AREA != 'USA')] <- 2

# now convert column to numeric since it is char
fp.t$AREA <- as.numeric(fp.t$AREA)

# check results
summary(fp.t$AREA)
```


2.  SERVICE_APPROACH - go for Service Approach S=1 or not S = 0
```{r, eval=FALSE}
# transform SERVICE_APPROACH: S = 1; !S = 0

fp.t$SERVICE_APPROACH[which(fp.data$SERVICE_APPROACH == 'S')] <- 1
fp.t$SERVICE_APPROACH[which(fp.data$SERVICE_APPROACH != 'S')] <- 0

# now convert column to numeric since it is char
fp.t$SERVICE_APPROACH <- as.numeric(fp.t$SERVICE_APPROACH)

# check results
summary(fp.t$SERVICE_APPROACH)

```


3.  CHANNEL_APPROACH - call it tier with second =1 and first =0
```{r, eval=FALSE}
# transform CHANNEL: rename to TIER with second = 1; first = 0

fp.t$CHANNEL_APPROACH[which(fp.data$CHANNEL_APPROACH == 'SECOND')] <- 1
fp.t$CHANNEL_APPROACH[which(fp.data$CHANNEL_APPROACH == 'FIRST')] <- 0

# now convert column to numeric since it is char
fp.t$CHANNEL_APPROACH <- as.numeric(fp.t$CHANNEL_APPROACH)

# check results
summary(fp.t$CHANNEL_APPROACH)

# now rename column to TIER
colnames(fp.t)[colnames(fp.t)=="CHANNEL_APPROACH"] <- "TIER"

head(fp.t$TIER)

```

4.  RENEWAL_FLAG - renewed = 1, not renewed = 0
```{r, eval=FALSE}
# transform RENEWAL_FLAG: RENEWED = 1; NOT RENEWED = 0

fp.t$RENEWAL_FLAG[which(fp.data$RENEWAL_FLAG == 'RENEWED')] <- 1
fp.t$RENEWAL_FLAG[which(fp.data$RENEWAL_FLAG == 'NOT RENEWED')] <- 0

# now convert column to numeric since it is char
fp.t$RENEWAL_FLAG <- as.numeric(fp.t$RENEWAL_FLAG)

# check results
summary(fp.t$RENEWAL_FLAG)

```


5.  Contract expiration date - ignore

__NO ACTION TAKEN__


6. CONTRACT_LENGTH - could go with annual = 1 and multi-year = 3 (most common value for each)

```{r, eval=FALSE}
# transform CONTRACT_LENGTH: ANNUAL = 1; MULTI-YEAR = 3

fp.t$CONTRACT_LENGTH[which(fp.data$CONTRACT_LENGTH == 'ANNUAL')] <- 1
fp.t$CONTRACT_LENGTH[which(fp.data$CONTRACT_LENGTH == 'MULTI-YEAR')] <- 3

# now convert column to numeric since it is char
fp.t$CONTRACT_LENGTH <- as.numeric(fp.t$CONTRACT_LENGTH)

# check results
summary(fp.t$CONTRACT_LENGTH)

```


7. Sales category - would separate into 3 new columns - ENT = 1, not ENT = 0, PS = 1, not PS=0, COM = 1, not COM = 0

```{r, eval=FALSE}
# transform SALES_CATEGORY - create 3 new columns COM, ENT, PS

# build COM column
# first check count of COM entries
sum(fp.data$SALES_CATEGORY == 'COM')

fp.t$COM <- 0
fp.t$COM[fp.data$SALES_CATEGORY == 'COM'] <- 1
# make sure new column matches original
sum(fp.t$ENT == 1)

# ---------------------------------
# build ENT column
sum(fp.data$SALES_CATEGORY == 'ENT')

fp.t$ENT <- 0
fp.t$ENT[fp.data$SALES_CATEGORY == 'ENT'] <- 1
sum(fp.t$ENT == 1)

# ---------------------------------
# build PS column
sum(fp.data$SALES_CATEGORY == 'PS')

fp.t$PS <- 0
fp.t$PS[fp.data$SALES_CATEGORY == 'PS'] <- 1
sum(fp.t$PS == 1)


```

8. Seller_unique ID - not relevant for binary conversation - too complex

__NO ACTION TAKEN__



9.  CONTRACT_VALUE_CATEGORY - going to bet it is collinear with cost and can be ignored - or could rank 1 to 5 lowest to highest buckets

```{r, eval=FALSE}
# transform CONTRACT_VALUE_CATEGORY: <10k = 0, 10K-25K = 1, etc..

fp.t$CONTRACT_VALUE_CATEGORY[which(fp.data$CONTRACT_VALUE_CATEGORY == '<10K')] <- 0
fp.t$CONTRACT_VALUE_CATEGORY[which(fp.data$CONTRACT_VALUE_CATEGORY == '10K-25K')] <- 1
fp.t$CONTRACT_VALUE_CATEGORY[which(fp.data$CONTRACT_VALUE_CATEGORY == '25K-50K')] <- 2
fp.t$CONTRACT_VALUE_CATEGORY[which(fp.data$CONTRACT_VALUE_CATEGORY == '50K-100K')] <- 3
fp.t$CONTRACT_VALUE_CATEGORY[which(fp.data$CONTRACT_VALUE_CATEGORY == '100K-250K')] <- 4
fp.t$CONTRACT_VALUE_CATEGORY[which(fp.data$CONTRACT_VALUE_CATEGORY == '>250K')] <- 5

# now convert column to numeric since it is char
fp.t$CONTRACT_VALUE_CATEGORY <- as.numeric(fp.t$CONTRACT_VALUE_CATEGORY)

# check results
summary(fp.t$CONTRACT_VALUE_CATEGORY)

```

10.  CONTRACT_LINE_CATEGORY- going to bet it is collinear with item count and can be ignored - or could rank 1 to 5 with lowest to highest category

```{r, eval=FALSE}
# transform CONTRACT_LINE_CATEGORY: VERY SMALL = 1, SMALL = 2, MEDIUM = 3, LARGE = 4, 
# VERY LARGE = 5

fp.t$CONTRACT_LINE_CATEGORY[which(fp.data$CONTRACT_LINE_CATEGORY == 'VERY SMALL')] <- 1
fp.t$CONTRACT_LINE_CATEGORY[which(fp.data$CONTRACT_LINE_CATEGORY == 'SMALL')] <- 2
fp.t$CONTRACT_LINE_CATEGORY[which(fp.data$CONTRACT_LINE_CATEGORY == 'MEDIUM')] <- 3
fp.t$CONTRACT_LINE_CATEGORY[which(fp.data$CONTRACT_LINE_CATEGORY == 'LARGE')] <- 4
fp.t$CONTRACT_LINE_CATEGORY[which(fp.data$CONTRACT_LINE_CATEGORY == 'VERY LARGE')] <- 5

# now convert column to numeric since it is char
fp.t$CONTRACT_LINE_CATEGORY <- as.numeric(fp.t$CONTRACT_LINE_CATEGORY)

# check results
summary(fp.t$CONTRACT_LINE_CATEGORY)

```

11.  DISCOUNT_CATEGORY - could go with no or some or call no 0, small 10, medium 20, large 30, very large 50

```{r, eval=FALSE}
# transform CONTRACT_LINE_CATEGORY: NO DISCOUNT = 0, SMALL = 10, MEDIUM = 20, LARGE = 30, 
# VERY LARGE = 50

fp.t$DISCOUNT_CATEGORY[which(fp.data$DISCOUNT_CATEGORY == 'NO DISCOUNT')] <- 0
fp.t$DISCOUNT_CATEGORY[which(fp.data$DISCOUNT_CATEGORY == 'MEDIUM')] <- 20
fp.t$DISCOUNT_CATEGORY[which(fp.data$DISCOUNT_CATEGORY == 'LARGE')] <- 30
fp.t$DISCOUNT_CATEGORY[which(fp.data$DISCOUNT_CATEGORY == 'VERY LARGE')] <- 50

# now convert column to numeric since it is char
fp.t$DISCOUNT_CATEGORY <- as.numeric(fp.t$DISCOUNT_CATEGORY)

# check results
summary(fp.t$DISCOUNT_CATEGORY)

```


11,  MULTIPLE_SERVICES = Yes = 1 and No = 0

```{r, eval=FALSE}
# transform MULTIPLE_SERVICES: YES = 1; NO = 0

fp.t$MULTIPLE_SERVICES[which(fp.data$MULTIPLE_SERVICES == 'YES')] <- 1
fp.t$MULTIPLE_SERVICES[which(fp.data$MULTIPLE_SERVICES == 'NO')] <- 0

# now convert column to numeric since it is char
fp.t$MULTIPLE_SERVICES <- as.numeric(fp.t$MULTIPLE_SERVICES)

# check results
summary(fp.t$MULTIPLE_SERVICES)

```


12.  SALES_STRATEGY - GCS = 1, not GCS = 0
```{r, eval=FALSE}
# transform SALES_STRATEGY: GCS = 1; NOT GCS = 0

fp.t$SALES_STRATEGY[which(fp.data$SALES_STRATEGY == 'GCS')] <- 1
fp.t$SALES_STRATEGY[which(fp.data$SALES_STRATEGY == 'NOT GCS')] <- 0

# now convert column to numeric since it is char
fp.t$SALES_STRATEGY <- as.numeric(fp.t$SALES_STRATEGY)

# check results
summary(fp.t$SALES_STRATEGY)

```


Now write transformed data set to a file
```{r, eval=FALSE}
write.csv(fp.t, file = "C:/SQLData/621-FP-Transformed-Data.csv", row.names = FALSE)
```


## Missing Data to be Imputed

The transformed data set can be seen here:

https://raw.githubusercontent.com/spsstudent15/2016-02-621-W2/master/621-FP-Transformed-Data.csv


